State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-07-12 StateOfPractice/Papers/SOP_MI_OSR/SOP_MI_OSR.tex
Ready for submission

\usepackage{booktabs}

\usepackage{amssymb}
\usepackage{amsmath,amsthm}

@@

\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage{pbox}

@@

\newcounter{pnum} %pain point number
\newcommand{\ppthepnum}{P`'\thepnum}
\newcommand{\ppref}[1]{P\ref{#1}}

\newcounter{qnum} %quality number
\newcommand{\qthepnum}{Q`'\theqnum}
\newcommand{\qref}[1]{Q\ref{#1}}

@@

%\pgfplotsset{compat=1.13}

@@

\author*[1]{\fnm{W.\ Spencer} \sur{Smith}}\email{smiths@mcmaster.ca}

\author[1]{\fnm{Ao} \sur{Dong}}

\author[1]{\fnm{Jacques} \sur{Carette}}\email{carette@mcmaster.ca}

\author[2]{\fnm{Michael} \sur{Noseworthy}}\email{nosewor@mcmaster.ca}

\affil*[1]{\orgdiv{Computing and Software Department}, \orgname{McMaster
University}, \orgaddress{\street{1280 Main Street West}, \city{Hamilton},
\postcode{L8S 4K1}, \state{Ontario}, \country{Canada}}}

\affil[2]{\orgdiv{Electrical and Computer Engineering Department},
\orgname{McMaster University}, \orgaddress{\street{1280 Main Street West},
\city{Hamilton}, \postcode{L8S 4K1}, \state{Ontario}, \country{Canada}}}

@@

We have a standard set of questions designed to assess the qualities of any
research software project~\cite{SmithEtAl2021, SmithAndMichalski2022}.  This has
been applied to MI software and Lattice Boltzmann Solvers \cite{SmithEtAl2024}.
This builds off prior work to assess the state of the practice for such domains
as Geographic Information Systems \cite{smith2018state}, Mesh Generators
\cite{smith2016state}, Seismology software \cite{Smith2018Seismology}, and
Statistical software for psychology \cite{smith2018statistical}.  We maintain
the previous constraint that the work load for measuring a given domain should
take around one person-month's worth of effort ($20$ working days at $8$
person-hours per day).

We identify a list of potential packages (through online searches) which is then
filtered and vetted by a domain expert. We aim for roughly $30$ packages. For
each remaining package, we measure its qualities by filling in a grading
template~\cite{SmithEtAl2021}.  This data is used to rank the projects with the
Analytic Hierarchy Process (AHP).  We summarize further details on the
interaction with the domain expert, software qualities, grading the software and
AHP below and in longer form in Smith et al.\
(2024)~\cite{SmithEtAl2024_MI_SOP}.

@@

The Domain Expert vets the proposed list because online resources can be
inaccurate.  The expert also vets the AHP ranking.  For the current assessment,
our Domain Expert is [Details of our domain expert removed for double-blind].
%%DOUBLE BLIND %our Domain Expert (and paper co-author) is Dr.\ Michael
%Noseworthy, Professor of Electrical and Computer Engineering at McMaster
%University, Co-Director of the McMaster School of Biomedical Engineering, and
%Director of Medical Imaging Physics and Engineering at St.\ Joseph's
%Healthcare, Hamilton, Ontario, Canada.

@@

We use an existing template~\cite{SmithEtAl2021} that is designed to measure the
aforementioned qualities. To stay within our given measurement time frame, each
package gets up to five hours of time.  Project developers can be contacted for
help regarding installation, if necessary, but we impose a cap of about two
hours on the installation process.  Figure~\ref{fg_grading_template_example}
shows an excerpt of the measurement spreadsheet.  The rows are the measures and
the columns correspond to the software packages.  [The full data is available on
Mendeley; link will be provided after refereeing.] 
% \cite{Dong2021-Data}
%provides the full set of measurement data.   %%DOUBLE BLIND

@@

The grader, after answering questions for each quality assigns an overall
score (between 1 and 10) based on the answers.  Several of the qualities use
the word ``surface'' to highlight that these particular qualities are a shallow
measure.  For example, usability is not measured using user studies.
Instead, we look for signs that the developers considered usability.
We use two freeware tools to collect repository related data:
\href{https://github.com/tomgi/git_stats}{GitStats} and
\href{https://github.com/boyter/scc}{Sloc Cloc and Code (scc)}.  Further
details on quality measurement are provided in Smith et al.\ (2024)
\cite{SmithEtAl2024_MI_SOP}.

@@

The measurement process used has previously been shown to be reasonably
reproducible.  \cite{SmithEtAl2016} reports grading five software
products by two reviewers. Their rankings were almost identical. As long as each
grader uses consistent definitions, the relative comparisons in the AHP results
will be consistent between graders.

@@

If the results of a study can be generalized to other situations, then the
study is externally valid \cite{RunesonAndHost2009}.  In other words, we cannot
generalize our results if the development of MI software is fundamentally
different from other research software.  Although there are differences, like
the importance of data privacy for MI data, we found the approach to developing
LBM software \cite{SmithEtAl2024} and MI software to be similar.  Except for
the domain specific aspects, the trends observed in the current study are
similar to that for other research software.

@@

We would like to thank Peter Michalski and Oluwaseun Owojaiye for fruitful
discussions on topics relevant to this paper.

>>>>

\usepackage{amssymb}
\usepackage{amsmath}

@@

\usepackage{multirow}
\usepackage{booktabs}
\usepackage{pbox}

@@

@@

@@

%% DOUBLE BLIND
% \author*[1]{\fnm{W.\ Spencer} \sur{Smith}}\email{smiths@mcmaster.ca}

% \author[1]{\fnm{Ao} \sur{Dong}}

% \author[1]{\fnm{Jacques} \sur{Carette}}\email{carette@mcmaster.ca}

% \author[2]{\fnm{Michael} \sur{Noseworthy}}\email{nosewor@mcmaster.ca}

% \affil*[1]{\orgdiv{Computing and Software Department}, \orgname{McMaster
% University}, \orgaddress{\street{1280 Main Street West}, \city{Hamilton},
% \postcode{L8S 4K1}, \state{Ontario}, \country{Canada}}}

% \affil[2]{\orgdiv{Electrical and Computer Engineering Department},
% \orgname{McMaster University}, \orgaddress{\street{1280 Main Street West},
% \city{Hamilton}, \postcode{L8S 4K1}, \state{Ontario}, \country{Canada}}}

@@

We have a standard set of questions designed to assess the qualities of any
research software project.  [blind review redacted details on the author's
previous studies using the same methodology.]
%% DOUBLE BLIND
% ~\cite{SmithEtAl2021, SmithAndMichalski2022}.  This has
% been applied to MI software and Lattice Boltzmann Solvers \cite{SmithEtAl2024}.
% This builds off prior work to assess the state of the practice for such domains
% as Geographic Information Systems \cite{smith2018state}, Mesh Generators
% \cite{smith2016state}, Seismology software \cite{Smith2018Seismology}, and
% Statistical software for psychology \cite{smith2018statistical}.  
We maintain
the previous constraint that the work load for measuring a given domain should
take around one person-month's worth of effort ($20$ working days at $8$
person-hours per day).

We identify a list of potential packages (through online searches) which is then
filtered and vetted by a domain expert. We aim for roughly $30$ packages. For
each remaining package, we measure its qualities by filling in a grading
template [citation redacted for double blind]. %%DOUBLE BLIND
%~\cite{SmithEtAl2021}.  
This data is used to rank the projects with the
Analytic Hierarchy Process (AHP).  We summarize further details on the
interaction with the domain expert, software qualities, grading the software and
AHP below and in longer form in [redacted for double blind].
%% DOUBLE BLIND
%Smith et al.\ (2024)~\cite{SmithEtAl2024_MI_SOP}.

@@

The Domain Expert vets the proposed list because online resources can be
inaccurate.  The expert also vets the AHP ranking.  For the current assessment,
our Domain Expert is [details of our domain expert removed for double-blind].
%%DOUBLE BLIND %our Domain Expert (and paper co-author) is Dr.\ Michael
%Noseworthy, Professor of Electrical and Computer Engineering at McMaster
%University, Co-Director of the McMaster School of Biomedical Engineering, and
%Director of Medical Imaging Physics and Engineering at St.\ Joseph's
%Healthcare, Hamilton, Ontario, Canada.

@@

We use an existing template [citation redacted]
%% DOUBLE BLIND
%~\cite{SmithEtAl2021} 
that is designed to measure the
aforementioned qualities. To stay within our given measurement time frame, each
package gets up to five hours of time.  Project developers can be contacted for
help regarding installation, if necessary, but we impose a cap of about two
hours on the installation process.  Figure~\ref{fg_grading_template_example}
shows an excerpt of the measurement spreadsheet.  The rows are the measures and
the columns correspond to the software packages.  [The full data is available on
Mendeley; link will be provided after refereeing.] 
% \cite{Dong2021-Data}
%provides the full set of measurement data.   %%DOUBLE BLIND

@@

The grader, after answering questions for each quality assigns an overall
score (between 1 and 10) based on the answers.  Several of the qualities use
the word ``surface'' to highlight that these particular qualities are a shallow
measure.  For example, usability is not measured using user studies.
Instead, we look for signs that the developers considered usability.
We use two freeware tools to collect repository related data:
\href{https://github.com/tomgi/git_stats}{GitStats} and
\href{https://github.com/boyter/scc}{Sloc Cloc and Code (scc)}.  Further
details on quality measurement are provided in [redacted for double blind].
%% DOUBLE BLIND
%Smith et al.\ (2024) \cite{SmithEtAl2024_MI_SOP}.

@@

The measurement process used has previously been shown to be reasonably
reproducible.  [citation redacted] 
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2016} 
reports grading five software
products by two reviewers. Their rankings were almost identical. As long as each
grader uses consistent definitions, the relative comparisons in the AHP results
will be consistent between graders.

@@

If the results of a study can be generalized to other situations, then the
study is externally valid \cite{RunesonAndHost2009}.  In other words, we cannot
generalize our results if the development of MI software is fundamentally
different from other research software.  Although there are differences, like
the importance of data privacy for MI data, we found the approach to developing
[other software (citation redacted)]
%% DOUBLE BLIND
%LBM software \cite{SmithEtAl2024} 
and MI software to be similar.  Except for
the domain specific aspects, the trends observed in the current study are
similar to that for other research software.

@@

[redacted for blind review]
%% DOUBLE BLIND
% We would like to thank Peter Michalski and Oluwaseun Owojaiye for fruitful
% discussions on topics relevant to this paper.
